### 2018-08-07 article.tex
minor edits
textrm to text
rewriting system example

\begin{definition}  \label{def:strict-two-cat}
  A \emph{strict} 2-category $\mathcal{C}$ is a category \emph{enriched over
    \textbf{Cat}}, the category of categories, \emph{viz:}
  \begin{itemize}
    \item A category $\mathcal{C}$, whose objects and arrows are known as
      \emph{0-cells} and \emph{1-cells} respectively.
    \item For every pair of objects $A$ and $B$, we have a
      category $\mathcal{C}(A, B)$, whose arrows are known as \emph{2-cells}.
    \item For \emph{2-cells} $\alpha : f \Rightarrow g$ and $\beta : g
      \Rightarrow h$ in $\mathcal{C}(A,B)$ their (vertical) composite is denoted
      $\beta \bullet \alpha : f \Rightarrow h$.
    \item For every 1-cell $f:A\to B$ we denote its identity 2-cell by
      $Id_f:f \Rightarrow f$.
    \item For all \emph{0-cells} $a$, $b$, $c$ a bifunctor $\circ:
      \mathcal{C}(b, c) \times \mathcal{C}(a, b) \to \mathcal{C}(a, c)$ called
      \emph{horizontal composition} of \emph{2-cells}.
  \end{itemize}
\REM{Need here the name for a 2-cat with all \emph{2-cells} invertible;  the
  term 2-groupoid refers to the case where both 1- and \emph{2-cells} are
  invertible.}
\end{definition}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        MONOIDAL 2-CATEGORY DEFINITION        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{definition}\label{def:monoidal-two-cat}
  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes)$ is a
  strict 2-category with a 2-bifunctor $\otimes: \mathcal{C} \times
  \mathcal{C} \to \mathcal{C}$.
  \TODO{Denis: fix this definition}
  \REM{What about the unit object?}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        INTERCHANGE LAW REMARK        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}[Interchange law]
\TODO{DENIS : please fix this statement.  THe second and third clauses
  are the same.!}
  Since $\circ$ and $\otimes$ are bifunctorial, we have the following
    interchange equations for all \emph{0-cells} $A, B \in \mathcal{C}$ and all
    \emph{2-cells} $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
  \begin{align}
    (\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet
      \gamma) \circ (\beta \bullet \delta) \\
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
      \gamma) \otimes (\beta \bullet \delta)
  \end{align}
  Since $\otimes$ is a 2-functor, we have for all \emph{0-cells} $A, B \in
    \mathcal{C}$ and all \emph{2-cells} $\alpha, \beta, \gamma, \delta \in
    \mathcal{C}(A, B)$
  \[
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) = (\alpha \bullet
      \gamma) \bullet (\beta \bullet \delta)
  \]
\end{remark}

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        REWRITING SYSTEM EXAMPLE        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{example}
  TODO rewrite sequence \& atomic rule
\end{example}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        HOMOMORPHISM PROPOSITION 2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proposition}
  The following category $\mathcal{P}(\mathbb{Q})$ is a $\mathbb{Q}$-category
  \begin{itemize}
    \item A set of objects which are the gates of $\mathcal{P}$
    \item For all objects $(f, g)$
  \end{itemize}
  \[
    \mathcal{E}(f, g) =
    \begin{cases}
      \textrm{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \\
      \textrm{otherwise } &\top
    \end{cases}
  \]
  With $\top \in \mathbb{Q}$ such that for all $x \in \mathbb{Q}$
  \[
    x \preceq \top
  \]

  \begin{proof}
    For all gates $(f, g)$, if it exists a unique transformation $f \Rightarrow
      g$, then
    \begin{align}
      \mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \\
      &= V(f) \oplus V(g)^{-1}
    \end{align}
    otherwise we already have $\mathcal{E}(f, g) = \top \in \mathbb{Q}$.
    For all gates $f$, we know that there is a identity transformation $f
      \Rightarrow f$ and
    \begin{align}
      \mathcal{E}(f, f) &= V(id_f) \\
      &= V(f \Rightarrow f)\\
      &= V(f) \oplus V(f)^{-1} \\
      &= e
    \end{align}
    Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f
      \Rightarrow g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the
      homomorphism $V$ give us
    \begin{align}
      \mathcal{E}(f, h) &= V(f \Rightarrow h) \\
      &= V(\alpha \circ \beta) \\
      &= V(\alpha) \oplus V(\beta) \\
      &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \\
      &= \mathcal{E}(g, h) \oplus \mathcal{E}(f, g)
    \end{align}
    If either $\alpha$ or $\beta$ doesn't exists in $\mathcal{P}$ then
    \[
      \mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
    \]
    become
    \[
      \begin{matrix}
        &\mathcal{E}(b, c) \oplus \top &\succeq \top \\
        \textrm{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \\
        \textrm{or} &\top \oplus \top &\succeq \top
      \end{matrix}
    \]
    which is still true.
  \end{proof}
\end{proposition}

>>>>

\begin{definition}  \label{def:strict-two-cat}
  A \emph{strict} 2-category $\mathcal{C}$ is a category \emph{enriched over
    \textbf{Cat}}, the category of categories, \emph{viz:}
  \begin{itemize}
    \item A category $\mathcal{C}$, whose objects and arrows are known as
      \emph{0-cells} and \emph{1-cells} respectively.
    \item For every pair of objects $A$ and $B$, we have a
      category $\mathcal{C}(A, B)$, whose arrows are known as \emph{2-cells}.
    \item For \emph{2-cells} $\alpha : f \Rightarrow g$ and $\beta : g
      \Rightarrow h$ in $\mathcal{C}(A,B)$ their (vertical) composite is denoted
      $\beta \bullet \alpha : f \Rightarrow h$.
    \item For every 1-cell $f:A\to B$ we denote its identity 2-cell by
      $Id_f:f \Rightarrow f$.
    \item For all \emph{0-cells} $A$, $B$, $C$ a bifunctor $\circ:
      \mathcal{C}(B, C) \times \mathcal{C}(A, B) \to \mathcal{C}(A, C)$ called
      \emph{horizontal composition} of \emph{2-cells}.
  \end{itemize}
\REM{Need here the name for a 2-cat with all \emph{2-cells} invertible;  the
  term 2-groupoid refers to the case where both \emph{1-} and \emph{2-cells} are
  invertible.}
\end{definition}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        MONOIDAL 2-CATEGORY DEFINITION        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{definition}\label{def:monoidal-two-cat}
  A \emph{strict} monoidal 2-category $(\mathcal{C}, \otimes, I)$ is a
  strict 2-category with a 2-bifunctor $\otimes: \mathcal{C} \times
  \mathcal{C} \to \mathcal{C}$ such that
  \begin{itemize}
    \item for all \emph{0-cells} $A$
      \[
        A \otimes I = A
      \]
    \item for all \emph{0-cells} $(A, B)$, $I$ is the monoidal unit of
    $\mathcal{C}(A, B)$
  \end{itemize}
  \TODO{Denis: fix this definition}
  \REM{What about the unit object?}
\end{definition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        INTERCHANGE LAW REMARK        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{remark}[Interchange law]
\TODO{DENIS : please fix this statement.  THe second and third clauses
  are the same.!}
  Since $\circ$ and $\otimes$ are bifunctorial, we have the following
    interchange equations for all \emph{0-cells} $A, B \in \mathcal{C}$ and all
    \emph{2-cells} $\alpha, \beta, \gamma, \delta \in \mathcal{C}(A, B)$
  \begin{align}
    (\alpha \circ \beta) \bullet (\gamma \circ \delta) &= (\alpha \bullet
      \gamma) \circ (\beta \bullet \delta) \\
    (\alpha \otimes \beta) \bullet (\gamma \otimes \delta) &= (\alpha \bullet
      \gamma) \otimes (\beta \bullet \delta)
  \end{align}
  Since $\otimes$ is a 2-functor, we have for all \emph{0-cells} $A, B \in
    \mathcal{C}$ and all \emph{2-cells} $\alpha, \beta, \gamma, \delta \in
    \mathcal{C}(A, B)$
  \[
    (\alpha \otimes \beta) \circ (\gamma \otimes \delta) = (\alpha \circ \gamma)
      \otimes (\beta \circ \delta)
  \]
\end{remark}

@@

As remarked earlier we don't include any equations in our notion of
circuits

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        REWRITING SYSTEM EXAMPLE        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{example}
  Let's take the previous quantum circuit example \ref{ex:quantum-circs} and add
    some \emph{noisy gates}, that is, gates which can failed with an $\epsilon$
    probability.
  \[
    \begin{array}{cccccccccc}
      \circZ{{\epsilon}} & \qquad  & \circX{{\epsilon}} & \qquad &
        \circbox{H_\epsilon} \\ 
      \vstrut{1.2em}
      Z_\epsilon : 1 \to 1 && X_\epsilon : 1 \to 1 && H_\epsilon : 1 \to 1
    \end{array}
  \]
  We form a 2-PROP by adding some rewriting rules (we omit the identies)
  \[
    \begin{array}{cccccccc}
      \circZ{{\epsilon}} & \Rightarrow & \circZ{{}} & \alpha : Z_\epsilon \to Z \\
      \vstrut{1.2em} \\
      \circX{{\epsilon}} & \Rightarrow & \circX{{}} & \beta : X_\epsilon \to X \\
      \vstrut{1.2em} \\
      \circbox{H_\epsilon} & \Rightarrow & \circH & \gamma : H_\epsilon \to H
    \end{array}
  \]
  Then, on the circuit
  \[
    \begin{array}{cccccccccc}
      \circbox{H_\epsilon} \\ 
      \vstrut{1.2em} \\
      \circZ{{\epsilon}}
    \end{array}
  \]
  a rewriting sequence can be $(id_{H_\epsilon} \otimes \alpha) \bullet (\beta
    \otimes id_Z)$
  \[
    \begin{array}{cccccccccc}
      \circbox{H_\epsilon} & \overset{id_{H_\epsilon}}{\Rightarrow} &
        \circbox{H_\epsilon} & \overset{\beta}{\Rightarrow} & \circH \\ 
      \vstrut{1.2em} \\
      \circZ{{\epsilon}} & \overset{\alpha}{\Rightarrow} & \circZ{{}} &
        \overset{id_Z}{\Rightarrow} & \circZ{{}} \\ 
    \end{array}
  \]
  Moreover this sequence is atomic.
\end{example}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%        HOMOMORPHISM PROPOSITION 2        %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{proposition}
  The following category $\mathcal{P}(\mathbb{Q})$ is a $\mathbb{Q}$-category
  \begin{itemize}
    \item A set of objects which are the gates of $\mathcal{P}$
    \item For all objects $(f, g)$
  \end{itemize}
  \[
    \mathcal{E}(f, g) =
    \begin{cases}
      \text{if } f \Rightarrow g \in \mathcal{P}, &V(f \Rightarrow g) \\
      \text{otherwise } &\top
    \end{cases}
  \]
  With $\top \in \mathbb{Q}$ such that for all $x \in \mathbb{Q}$
  \[
    x \preceq \top
  \]

  \begin{proof}
    For all gates $(f, g)$, if it exists a unique transformation $f \Rightarrow
      g$, then
    \begin{align}
      \mathcal{E}(f, g) &= V(f \Rightarrow g) \in \mathbb{Q} \\
      &= V(f) \oplus V(g)^{-1}
    \end{align}
    otherwise we already have $\mathcal{E}(f, g) = \top \in \mathbb{Q}$.
    For all gates $f$, we know that there is a identity transformation $f
      \Rightarrow f$ and
    \begin{align}
      \mathcal{E}(f, f) &= V(id_f) \\
      &= V(f \Rightarrow f)\\
      &= V(f) \oplus V(f)^{-1} \\
      &= e
    \end{align}
    Finally, for all gates $(f, g, h)$ such that there exists $\alpha: f
      \Rightarrow g$ and $\beta: g \Rightarrow h$ in $\mathcal{P}$, then the
      homomorphism $V$ give us
    \begin{align}
      \mathcal{E}(f, h) &= V(f \Rightarrow h) \\
      &= V(\alpha \circ \beta) \\
      &= V(\alpha) \oplus V(\beta) \\
      &= V(g \Rightarrow h) \oplus V(f \Rightarrow g) \\
      &= \mathcal{E}(g, h) \oplus \mathcal{E}(f, g)
    \end{align}
    If either $\alpha$ or $\beta$ doesn't exists in $\mathcal{P}$ then
    \[
      \mathcal{E}(b, c) \oplus \mathcal{E}(a, b) \succeq \mathcal{E}(a, c)
    \]
    become
    \[
      \begin{matrix}
        &\mathcal{E}(b, c) \oplus \top &\succeq \top \\
        \text{or} &\top \oplus \mathcal{E}(a, b) &\succeq \top \\
        \text{or} &\top \oplus \top &\succeq \top
      \end{matrix}
    \]
    which is still true.
  \end{proof}
\end{proposition}
